---
title: "Standardized Network Analysis-TD Comparison"
author: "Stanley West"
format: 
  revealjs:
    embed-resources: true
    self-contained-math: true
    scrollable: true
    smaller: true
    fig-width: 24
    fig-height: 12
    fig-align: center
    
    
editor: 
  markdown: 
    wrap: 72
---

# The Data {.smaller .scrollable}

-   Up to this point, individual lexical networks have been constructed
    for each child, with edges drawn according to forward association
    (i.e., cue -\> response) relationships via child-oriented word
    association norms ([Cox & Haebig,
    2023](https://link.springer.com/content/pdf/10.3758/s13428-022-01790-y.pdf)),
    or noun feature norms ([Borovsky, Peters, Cox, &
    McRae](https://link.springer.com/article/10.3758/s13428-023-02242-x)),
    both with all features, and with features filtered to only include a
    toddler accessibility rating (TAR) greater than 4.
    -   Network statistics including in-degree (number of incoming
        connections; median and average), global clustering coefficient
        (ratio of number of triangles to the number of possible
        triangles, average), average shortest path length (ASPL, average
        edge distance between nodes) were computed for each child.
    -   These values were standardized according to 1,000 individualized
        pseudo-random acquisition networks (RAN) for each child where
        the number of nodes as well as the number of nouns, verbs, and
        other words were matched to the child's vocabulary. These random
        networks were sampled from a main network of all the words
        available on the Words and Sentences and Words and Gestures
        forms, respectively. That is, a child's RAN network sample
        consisted of only the words that appear on the form from which
        their actual vocabulary network was contrived.

```{r,echo = FALSE}
library(tidyverse)
library(ggplot2)
library(purrr)
library(MatchIt)
library(ggh4x)
library(plotly)
library(ggraph)

d_netstats <- read_rds("data/formatted_metadata_crc.rds") %>%
  mutate(nProduced = nproduced,
         netBuild = network,
         metric_plot = factor(netstat, levels = c("degree","clust","dist"),
                              labels = c("Median Indegree", "Clustering Coefficient", "ASPL")),
         z = vocab_RANz) 

matched_pps <- read_rds("data/matched_groups_single_timepoint.rds")

load("data/graph_full_nounNet.Rdata")
load("data/child_oriented_graph.Rdata")

```

## Overall Statistic Trajectories Over Vocabulary Size and Age {.smaller .scrollable}

-   Looking at the distributions of raw and RAN scores across the
    different network constructions.

    -   'mean' and 'sd' in the bottom plot indicate averages and
        standard deviations of the RAN simulated statistics.

    -   The deviation of RAN mean estimations are due to the separate
        RAN simulations for the WS and WG forms.

```{r}

ggplot(d_netstats %>% 
         ungroup() %>%
         select(nProduced,z,netBuild,metric_plot) %>%
         unique(), aes(x= nProduced, y = z, color = netBuild))+
  geom_point()+
  geom_smooth()+
  theme_bw(base_size = 24)+
  facet_wrap(~metric_plot)+
  geom_hline(yintercept = 0)

ggplot(d_netstats %>% 
         filter(type == "vocab"), aes(x = nProduced, y = value))+
  geom_point()+
  geom_smooth(method = "loess")+
  facet_grid2(netBuild~metric_plot,scales = "free", independent = "all")+
  theme_bw(base_size = 24)+
  labs(title = "True Scores Across Vocabuary")

ggplot(d_netstats %>% 
         filter(type == "randomM"), aes(x = nProduced, y = value))+
  geom_point()+
  geom_smooth(method = "loess")+
  facet_grid2(netBuild~metric_plot,scales = "free", independent = "all")+
  theme_bw(base_size = 24)+
  labs(title = "RAN Mean Scores Across Vocabuary")

ggplot(d_netstats %>% 
         filter(type == "randomS"), aes(x = nProduced, y = value))+
  geom_point()+
  geom_smooth(method = "loess")+
  facet_grid2(netBuild~metric_plot,scales = "free", independent = "all")+
  theme_bw(base_size = 24)+
  labs(title = "RAN SD Scores Across Vocabuary")


ggplot(d_netstats, aes(x = nProduced, y = z))+
  geom_point()+
  geom_smooth(method = "loess")+
  facet_grid2(netBuild~metric_plot,scales = "free",independent = "all")+
  theme_bw(base_size = 24)+
  labs(title = "Standardized Scores across vocabuary")


# ggplot(d_netstats %>%
#          group_by(subjectkey_intAge,metric,netBuild) %>%
#          mutate(RAN_dist = (value[stat == "truestat"] - value[stat =="mean"])), aes(x = nProduced,y = RAN_dist, fill = netBuild))+
#   geom_col(position = "dodge")+
#   facet_wrap2(~metric_plot,scales = "free")+
#   theme_bw(base_size = 24)+
#   labs(y = "true stat - mean RAN stat")
# 
# ggsave("Figures/RAN_distance_plot.png")


ggplot(d_netstats%>% 
         ungroup() %>%
         select(nProduced,value,type,netBuild,metric_plot) %>%
         unique(), aes(x = nProduced, y = value, color = type))+
  geom_point(size = 2, alpha = .5)+
  theme_bw(base_size = 24)+
  facet_grid2(netBuild~metric_plot, scales = "free", independent = "all")
ggsave("Figures/network_stats_comparison.png")


```

<!-- ## Plotting unmatched statistic distributions over production {.smaller .scrollable .hidden} -->

<!-- -   Distributions across the unmatched samples look similar, especially -->
<!--     at smaller vocabularies. -->
<!-- -   However, this is still collapsed across forms. -->

<!-- ```{r} -->

<!-- ggplot(d_netstats %>%  -->
<!--          filter(form == "WS"), aes(x = nProduced, y = z, color = group))+ -->
<!--   facet_grid2(netBuild~metric_plot,scales = "free",independent = "all")+ -->
<!--   geom_smooth(method = "lm", -->
<!--               formula = y~poly(x,3))+ -->
<!--   theme_bw(base_size = 24)+ -->
<!--   labs(title = "Unmatched Statistic Distrubution Over Production - Words & Sentences", -->
<!--        x = "Productive Vocabulary", -->
<!--        y = "Z-scored Network Statistic") -->

<!-- ggplot(d_netstats %>%  -->
<!--          filter(form == "WG"), aes(x = nProduced, y = z, color = group))+ -->
<!--   geom_point() + -->
<!--   facet_grid2(netBuild~metric_plot,scales = "free",independent = "all")+ -->
<!--   geom_smooth(method = "lm", -->
<!--               formula = y~poly(x,3))+ -->
<!--   theme_bw(base_size = 24)+ -->
<!--   labs(title = "Unmatched Statistic Distrubution Over Production - Words & Gestures", -->
<!--        x = "Productive Vocabulary", -->
<!--        y = "Z-scored Network Statistic") -->



<!-- ggplot(d_netstats , aes(x = group, y = z, color = group))+ -->
<!--   geom_violin() + -->
<!--   geom_boxplot(width = 0.2, fill = "lightgrey")+ -->
<!--   facet_grid2(netBuild~metric_plot,scales = "free",independent = "all")+ -->
<!--   theme_bw(base_size = 24)+ -->
<!--   labs(title = "Unmatched Statistic Boxplots Over Group ", -->
<!--        x = "Group", -->
<!--        y = "Z-scored Network Statistic") -->

<!-- ``` -->

<!-- -   The middle group does not have any child with production past 100 -->
<!--     for the words and gestures form, making it hard to compare to the -->
<!--     other groups, so we removed this group. -->

## First Administrations - Both Forms

-   Including both WS and WG, but only taking the first administration:
    -   Removing 4 largest vocabularies in the delay group to help with
        matching
-   nProduced is the number of nouns produced.

```{r}


mdl_split <- d_netstats %>%
  filter(group == "mdl",
         (nProduced <=110 & interview_age >= 72)) %>%
  mutate(group_two = "delay",
         group_two = as.factor(group_two))


d_netstats_first <- d_netstats %>%
  filter(!group == "mdl",nProduced >= 3) %>%
  mutate(group_two = factor(ifelse(group == "yng","no_delay","delay"), levels = c("delay","no_delay"), labels = c("delay","no_delay"))) %>%
  filter(type == "vocab") %>%
  rbind(mdl_split)%>%
  ungroup() %>%
  filter(!(group_two == "delay" & nProduced >= 250),
         !(group_two == "no_delay" & nProduced >= 250))


d_netstats_first %>%
  select(subjectkey,group_two,nProduced, interview_age,netBuild) %>%
  unique() %>%
  group_by(group_two,netBuild) %>%
  summarize(n = n(),
            mean_nProduced = mean(nProduced),
            sd_nProduced = sd(nProduced),
            mean_age = mean(interview_age),
            sd_age = sd(interview_age)) %>%
  as.data.frame()


d_netstats_first$group_two <- relevel(d_netstats_first$group_two, ref = "no_delay")

```

## Matching between delay, no delay, and TD groups.

-   3-way ANOVA revealed no differences in number of nouns produced between the TD and two ASD groups $F(2,485) = 0.1, p = 0.905$.

```{r}

noun_assoc_split <- split(d_netstats_first,d_netstats_first$netBuild)


match_noun_assoc <- map(noun_assoc_split, function(x){
  matchit_all <- x %>%
    select(subjectkey, interview_age,nProduced, group_two) %>%
    unique() %>%
    filter(nProduced <= max(x$nProduced[x$group_two == "delay"]+20)) %>%
    matchit(group_two ~ nProduced, data = ., method = "optimal", distance = "glm"  ,ratio = 2)
  match_df <- x %>%
     filter(subjectkey %in% match.data(matchit_all)$subjectkey) 
  return(match_df)
})

d_netstats_TD <- read_rds("data/formatted_metadata_TD_crc.rds") %>%
  mutate(nProduced = nproduced,netBuild = network,
         metric_plot = factor(netstat, levels = c("degree","clust","dist"),
                              labels = c("Median Indegree", "Clustering Coefficient", "ASPL")),
         z = vocab_RANz) %>%
  mutate(group_two = "TD",
         group = "TD",
         interview_date = as.Date(interview_date)) %>%
  split(.,.$netBuild)

combined_list <- list(assoc = rbind(match_noun_assoc$assoc, d_netstats_TD$assoc),
                      feat = rbind(match_noun_assoc$feat,d_netstats_TD$feat))

map(combined_list, function(x){
  d <- x %>%
    select(subjectkey,nProduced,group_two) %>%
    unique() 
  return(summary(aov(nProduced~group_two, data = d)))
})[1]


```

-   Keeping both WS and WG forms, we match the no-delay group to the 41
    delayed children, allowing the no-delay group to be 2x bigger.

```{r, echo =  FALSE,fig.width=48, fig.height=12, fig.fullwidth=TRUE}

map(combined_list, function(x){
  x %>%
    select(subjectkey,interview_age,nProduced,group_two) %>%
    unique() %>%
    group_by(group_two) %>%
    summarize(n = n(),
              mean_age = mean(interview_age),
              mean_produced = mean(nProduced),
              sd_age = sd(interview_age),
              sd_produced = sd(nProduced)) %>%
    as.data.frame()
})[1]

map(combined_list, ~
        ggplot(.x, aes(x = nProduced, color = group_two))+
          geom_density(size = 2)+
          theme_bw(base_size = 52)+
          labs(title = paste0("Matching between groups ", unique(.x$netBuild))))[1]

map(combined_list, ~
        ggplot(.x, aes(x = nProduced, , y = interview_age,color = group_two))+
          geom_point(size = 13)+
          theme_bw(base_size = 52)+
          labs(title = paste0("Matching between groups ", unique(.x$netBuild))))[1]


```

## Run Linear Models {.smaller .scrollable}

-   Number of words produced was re-factored to account for third-order
    polynomial trends in the network statistics

-   Nested model comparisons revealed that only including linear and
    quadratic trends leads to best model fit.

-   Linear models were run regressing
    $network_{statistic} \sim (linear_{nproduced} + quadratic_{nproduced})  * group$
    
-   The reference group is set to the typically developing group. 

-   Noun Features (TAR filtered; interpreted by Chris Cox):


    -  Median indegree: There is a significant linear trend in median indegree, as well as significant mean differences between the TD and both the delay and no-delay groups. A signficant negative interaction between group and linear trends indicate that the TD group is increasing faster than both delay and no-delay groups. 

    - Clustering Coefficient:  There are no trends, linear or quadratic, for the clustering coefficient, but there is a main effect of group whereby the delayed group tends to have higher clustering coefficients on average.

    -  ASPL: There is a negative linear trend (and no quadratic trend), as well as mean differences between both ASD groups (delay and no-delay) and the TD group. There is also a significant linear by group interaction for both groups indicating that the TD group is achieving low path lengths faster. Finally, there is a significant quadratic by group interaction for the no delay group, indicated by a late increase in path length around 200 nouns.

-   Noun-only Associations (interpreted by Chris Cox):

    -  Median Indegree: There is a positive linear and a negative quadradic trend, as well as a significant negaitive group difference between both ASD groups and the TD group. There are negative linear and positive quadratic by no-delay group interactions, captured by rapid decreases in indegree early on for the no-delay group, whereas the TD group has steady increases over noun acquisition.

    - Clustering Coefficient: There are positive linear and negative quadratic trends as well as significant group differences between the TD and both ASD groups. There are also significant negative linear and positive quadratic by no-delay group interactions, suggesting steeper decrease in clustering for the no-delay group, with steady increases that plateau in the TD group. There was finally a significant negaitve linear by delay group interaction, again with this group decreasing faster than the TD group.

    - ASPL: There are negative linear and positive quadratic trends in ASPL, with no significant group differences. There are also positive linear by delay/no-delay interactions, suggesting quicker decrease in ASPL for the TD group early in noun acquisition. There was also a negative quadratic by delay group interaction, suggesting that the TD group has quicker increase in ASPL later in noun learning.

```{r, echo = TRUE}

nProduced_poly_all <- combined_list %>%
  map(~ .x %>%
        ungroup() %>%
        select(nProduced) %>%
        unique() %>%
        cbind(poly(.$nProduced,3)) %>%
        rename(linear = "1",
               quadratic = "2",
               cubic = "3"))

matched_df_poly_all <- map2(combined_list,nProduced_poly_all, function(x,y){
                              x %>%
                              left_join(y) %>%
                              filter(type == "vocab")
}) 
 

d_splits_all <- matched_df_poly_all %>%
  map(~ split(.x,.x$metric_plot))

map(d_splits_all, function(x){
  map(x, function(y){
     m1 <- lm(z~linear*group_two, data = y)
     m2 <- lm(z~(linear+quadratic)*group_two, data = y)
     m3 <- lm(z~(linear+quadratic+cubic)*group_two, data = y )
  return(anova(m1,m2,m3))
  })
})

wt <- ifelse(d_splits_all$assoc$`Median Indegree`$group_two == 'delay',2,1)

x<- map(d_splits_all, function(x){
    map(x, function(d){ 
      d$group_two <- relevel(d$group_two, ref = "TD")
          lm(z ~ (linear+quadratic)*group_two, data = d,weights = wt)
      })
})

map(x, ~map(.x, function(d){summary(d)}))




```

## Plot the Trends

```{r}

map(matched_df_poly_all, ~ ggplot(.x %>% 
         filter(type == "vocab"), aes(x = nProduced, y = z, color = group_two))+
  #geom_point()+
  theme_bw(base_size = 24) +
  geom_smooth(method = "lm",
              formula = y ~ poly(x,3))+
  facet_grid(~metric_plot)+
  labs(title = "Network Statistic Trajectory by Group",
       subtitle = paste("Network Construction:",unique(.x$netBuild))))




```

